---
title: "data_exploration"
author: "Binyam Yilma"
date: "3/14/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(caret) 
library(splines)
library(mgcv) #mainly for the GAM models
library(pdp) #partial dependent plot
library(earth) #for the MARS models
library(tidyverse)
library(ggplot2)
library(summarytools)


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
 ggplot2.continous.colour = "viridis",
 ggplot2.continous.fill = "viridis"
)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "100%",
  warning = F,
  message = F
)


scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(2021)
```

Import data
```{r}
data <- read_csv("./data/us_housing_archive/data.csv")


#Number of distinct houses
data %>% 
  distinct(street, city) %>% 
  count() %>% 
  knitr::kable(caption = "n_distinct_houses")
```


1 - Data study
2 - Exploratory data analysis
3 - Feature engineering
4 - Prediction


Data Study
```{r}
#Visualize missing data. 
visdat::vis_miss(data) 
```
There are no missing data. 


```{r}
# Data was recorded from May 05, 2014 until July 07, 2014 - over 70 days. 
data %>% 
  distinct(date) %>% 
  summarise(min = min(date),
            max = max(date),
            range = n_distinct(date)
            ) %>% 
  knitr::kable(caption = "Range of dates ")

```

Feature plot of variables against the outcome - price
```{r}
featurePlot(x = data[, 3:14], 
            y = data$price, 
            plot = "scatter", 
            span = .5, 
            labels = c("Predictors","Y"),
            type = c("p", "smooth"))
```

To do:

1 - Drop `street`, `state`, `country`
2 - Transform `yr_built` & `yr_renovated`, into `house_age` & `yr_since_renovation`
   - Recode `0` (w/c means house has not been renovated) in `yr_since_renovation` to be equals to the `house_age` 
3 - Drop `Date`
4 - Split `state zip` variable into two - drop `state`, keep `zip.`
5 - Build a data library descrbing what each variables are and how they're coded. 
6 - Aply box-cox transformation on the outcome variable
7 - Turn categorical variables - `waterfront`, `view`, `condition` -  in `factors` 


Plotting the predictor outcome. 

```{r, include=F}
data %>% 
  distinct(statezip) 
  
data %>% 
  distinct(yr_built) 

data %>% 
  distinct(yr_built, yr_renovated)

data %>% 
  distinct(view)
```


